An Automatic Verification Framework for Software Configurations

### 2016-05-10 OSDI/learner.tex
slight algo edits
add algo description

The first component learns two types of rules: rules that are inferred  using typed templates and 
rules which we call \emph{untyped specifications}. Those are rules
that must hold over multiple parts of a configuration file. An example of an
untyped specification is the ordering or missing entry constraint.

@@

\begin{algorithm}[t]
\caption{Construct Rules}
\label{alg:crules}
\begin{footnotesize}
  \begin{algorithmic}[1]
 \State \textbf{constructRules}($e, z, opt, S$):\\
    Input: entry $e$ from intermediate representation\\
     \qquad \enspace \enspace type or untyped specification $z$\\
     \qquad \enspace \enspace option $opt$, set to 1 if template\\
     \qquad \qquad \enspace \enspace \enspace \enspace \qquad or 0 if untyped spec\\
     \qquad \enspace \enspace set of entries $S$\\
     \State $R = \{\}$
     \State Set $j$ as an iteration variable.
     \If{opt == 1}
        \State iterate the following sequence \textbf{L} over all elements of $templates(z)$ 
        \State \ie, $j \in templates(z)$
     \ElsIf{opt == 0}
        \State $j = z$
     \EndIf
     \State Beginning of \textbf{L}: 
     \State Let $n$ be the arity of $j$.
      \State Let $P$ be the set of permutations of distinct entries from $S$, of size $n-1$.
      \For{each $p \in P$}
        \State Let $p$ be represented by $(e_1, e_2, \ldots, e_{n-1})$.
         \For{$i = 1:n$}
         \State Create a $n$-tuple by placing $e$ at the $i$-th position and filling the remaining entries with $p$, in order. Let this $n$-tuple be $v = (e_1, e_2, \ldots, e_i, \ldots, e_{n-1})$.
         \State $r = \textrm{constructRule}(v)$
         \If{$r$ is satisfied}
                $R = R \cup r$
         \EndIf
        \EndFor
     \EndFor
     \State End of \textbf{L}. 
     \State return $R$
  \end{algorithmic}
\end{footnotesize}
\end{algorithm}

%\begin{algorithm}
%\caption{constructRule}
%  \begin{algorithmic}[1]
%  \State \textbf{constructRule}($u$):\\
%  Input: a tuple of entries, $u$, denoted by $(e_1, e_2, \ldots, e_i, \ldots, e_{n-1})$ \\
%    \State return $t(k(e_1), k(e_2), \ldots, k(e_i), \ldots, k(e_{n-1}))$.
%  \end{algorithmic}
%\end{algorithm}

%each entry is listed as (g,k,v,typelist)

@@

Note that the formalism relies on Algorithm~\ref{alg:plearn}. For brevity, Algorithm 1 only considers
templates without mixed types; that is, the arguments of the template are all of the same type. This is
demonstrated via the filtration step, which filters rules in a file that are possibly of the same type.  
Adding support for mixed types would be handled in a branching statement in Algorithm 1.
Note that in Algorithm~\ref{alg:crules}, constructRule refers to the creation of a rule from a tuple of entries, $u = (e_1, e_2, \ldots, e_i, \ldots, e_{n-1})$.
Then the rule creation is formally represented as $t(k(e_1), k(e_2), \ldots, k(e_i), \ldots, k(e_{n-1}))$; in order to 
determine if the rule satisfied, the relation associated with $t(v(e_1), v(e_2), \ldots, v(e_i), \ldots, v(e_{n-1}))$ is evaluated for truth.

>>>>

The first component learns two types of rules: rules that are inferred using templates associated with types 
and rules which we call \emph{untyped specifications}. An example of an
untyped specification is the ordering or missing entry constraint.
%Both are rules that must hold over multiple parts of a configuration file.

@@

\begin{algorithm}[t]
\caption{Construct Rules}
\label{alg:crules}
\begin{footnotesize}
  \begin{algorithmic}[1]
 \State \textbf{constructRules}($e, z, opt, S$):\\
    Input: entry $e$ from intermediate representation\\
     \qquad \enspace \enspace type or untyped specification $z$\\
     \qquad \enspace \enspace option $opt$, set to 1 if template\\
     \qquad \qquad \enspace \enspace \enspace \enspace \qquad or 0 if untyped spec\\
     \qquad \enspace \enspace set of entries $S$\\
     \State $R = \{\}$
     \State Set $j$ as an iteration variable.
     \If{opt == 1}
        \State run the following sequence \textbf{L} over all elements of $templates(z)$ 
        \State \ie, iterate $j$ over $templates(z)$
     \ElsIf{opt == 0}
        \State $j = z$
     \EndIf
     \State Beginning of \textbf{L}: 
     \State Let $n$ be the arity of $j$.
      \State Let $P$ be the set of permutations of distinct entries from $S$, of size $n-1$.
      \For{each $p \in P$}
        \State Let $p$ be represented by $(e_1, e_2, \ldots, e_{n-1})$.
         \For{$i = 1:n$}
         \State Create a $n$-tuple by placing $e$ at the $i$-th position and filling the remaining entries with $p$, in order. Let this $n$-tuple be $m = (e_1, e_2, \ldots, e_i, \ldots, e_{n-1})$.
         \State $r = \textrm{constructRule}(m)$
         \If{$r$ is satisfied}
                $R = R \cup r$
         \EndIf
        \EndFor
     \EndFor
     \State End of \textbf{L}. 
     \State return $R$
  \end{algorithmic}
\end{footnotesize}
\end{algorithm}

@@

Note that the formalism relies on Algorithm~\ref{alg:plearn}. The algorithm considers both typed templates and
untyped specifications. As a similar iteration logic is used for both, an $opt$ flag distinguishes between the two situations.
constructRule in Algorithm~\ref{alg:crules} refers to the creation of a rule from a tuple of entries, 
$m = (e_1, e_2, \ldots, e_i, \ldots, e_{n-1})$.
The rule creation is formally represented as $m(k(e_1), k(e_2), \ldots, k(e_i), \ldots, k(e_{n-1}))$. 
In order to determine if the rule satisfied, the relation associated with $m(v(e_1), v(e_2), \ldots, v(e_i), \ldots, v(e_{n-1}))$ is evaluated for truth.
The rule can be template-generated, for instance {\tt max\_connections} > {\tt mysql.max\_persistent}, or
it can be untyped-specification-generated rule, for instance {\tt recode.so} must come before 
{\tt mysql.so}.

For typed templates, we first filter and examine entries associated with a particular type, and then check if
the template is satisfied over appropriately typed argument entries in the rest of the file. We do something similar for
untyped specifications.
